---
title: |
  | Programming in Base R
  | Matt Bray
format: html
editor: visual
---

# Task 1 Step 1

## Create Pre Dose and Post Dose Blood Pressure Vector with the Same Names for each Subject

```{r}
preTreatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

postTreatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)

names(preTreatment) <- paste("Subject", 1:20, sep = "_")
names(postTreatment) <- paste("Subject", 1:20, sep = "_")
```

# Task 1 Step 3

## Calculate the change in blood pressure for each patient

```{r}
deltaBp <- preTreatment-postTreatment
```

# Task 1 Step 4

## Calculate the average decrease in blood pressure across all patients.

```{r}
avgDeltabp <- mean(deltaBp)
avgDeltabp
```

# Task 1 Step 5

## Determine which patients experienced a decrease in blood pressure after treatment.

```{r}
decrPatient <- deltaBp[which(deltaBp>0)]
decrPatient
```

# Task 1 Step 6

## Subset patients with decrease in blood pressure after treatment.

```{r}
d <- deltaBp[deltaBp>0]
d
avGd <- mean(d)
avGd
```

# Task 2 Step 1

## Create a data frame object with four columns.

```{r}
df <- data.frame(preTreatment, postTreatment, deltaBp)
df

```

# Task 2 Step 2

## Return only rows of the data frame where the deltaBp column is negative

```{r}
incrPatient <- df[df$deltaBp<0,]
incrPatient
```

# Task 2 Step 3

## Add a new column to the data frame corresponding to TRUE if preTreatment is less than 120 and FALSE if it is not.

```{r}
df$lessThan<-df$postTreatment<120
df
```

### Print DF

```{r}
knitr::kable(df)
```

# Task 3

## Create Placebo Dataset

```{r}
placeboPre<-c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placeboPost<-c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
names(placeboPre) <- paste("Subject", 1:10, sep = "_")
names(placeboPost) <- paste("Subject", 1:10, sep = "_")
deltaPlacebo<-placeboPre-placeboPost
```

# Task 3 Step 1

## Create placebo dataframe

```{r}
dfPlacebo<-data.frame(placeboPre, placeboPost, deltaPlacebo)
dfPlacebo$less120<-dfPlacebo$placeboPost<120
dfPlacebo
```

### Print dfPlacebo

```{r}
knitr::kable(dfPlacebo)
```

# Task 3 Step 2

## Create list of two elements...Treatment and Placebo Data frames

```{r}
list <- list("df"=df,"dfPlacebo"=dfPlacebo)
```

# Task 3 Step 3

## Access first list element in three ways.

```{r}
list[1]
list$df$preTreatment
list(names(df))
```

# Task 3 Step 4

## Use single line of code to access the placeboPre column of the placebo dataframe.

```{r}
list[[2]][1]
```
